

\subsection*{Outline: Detailed Work Plan for Conducting the Management Strategy Evaluation}
\label{sub:outline}

This section outlines, in tabular form, a schedule for the work required to conduct the 2014 and 2015 assessments and then 
to evaluate a management procedure using an operating model for Atlantic bluefin. A detailed work plan, based on this schedule, 
will be developed for presentation at the SCRS plenary. Following endorsement by the SCRS, a budget will be proposed for 
presentation to the Commission. To implement the operating model, it is essential that contracts for providing external 
support for a number of years are awarded. It is also essential to establish a core steering group to oversee the work.

\subsubsection*{2013}

\begin{enumerate}
\item Discussion of alternative mixing structures in broad terms SCRS paper with key contributors
\item Clarification of standard inputs to standard separate west/east assessments.
Use ICCAT meeting to facilitate with those most familiar with data (Terms of Reference document)
Table of information available
\item Clarification on data availability for mixing and stock structure related data for more complex stock assessments.
Genetic, microconstituents, tags (archival, conventional, other)
\item Identification of major sensitivities for both separate and mixed stock assessments (e.g., M, fecundity schedule, SRR and alternative mechanism of population regulation)
\item Use Risk assessment paper on qualitative identification of uncertainty (written under GBYP modelling contract) to inform OM scenarios, i.e., SCRS paper with key contributors.
\item Identification of those who will be taking both assessment approaches further forward
Consistent, core group over a multi-year timeline
\item Support capacity development for conduct, understanding and use of MSE in adoption of Harvest Control Rules for the Atlantic bluefin fisheries through:
ICES/ICCAT MSE training in (Dec. 2013) to facilitate capacity building for CPC scientific delegations; 
Take advantage of GEF/FAO Areas Beyond National Jurisdiction Tuna Program funds intended to accelerate the joint tuna RFMO working group for MSE development and management /stakeholder / science (jargon-free) dialogue.
Conduct a ‘side event’(SCRS Chairman to co-ordinate) at the 2013 Commission meeting open to CPCs and stakeholder groups, drawing upon the experience at CCSBT to initiate the management / science/ stakeholder dialogue. 
\end{enumerate}


\subsubsection*{2014} 

For bluefin session
\begin{enumerate}
\item Eastern assessment update 
Post-bluefin; ideally follows bluefin session
\item Review of updated separate assessment approaches
\item Review of initial mixed stock models and refinement of alternative mixing structure scenarios
\item Tool for visualizing movement
\item Meeting including stakeholders (finalise at 2013 Commission meeting)
\end{enumerate}

\subsubsection*{2015}


For bluefin session
\begin{enumerate}
\item Data guillotine – finalisation of data to be considered for operating model development for MP testing (back reference); includes data update for assessments for that year.
\item Updated catch limit advice for both East and West based on revised separate assessments, and possibly also mixed stock models
\end{enumerate}

\subsubsection*{2016}

For separate meeting (Jan-Feb 2016)
\begin{enumerate}
\item Agreement on operating model specifications (conditioning)
\item Agreement on data for use in MPs
Subset of OM data
\item Initial agreement on objectives and performance statistics (other stakeholders need to be included in these discussions)
\item Specifications and schedule for coding
Early in year
\item Circulation of code to allow alternative MP developers to plug-in and test their candidate MPs
Bluefin session
\item Refinement of MP testing procedures
\item Interaction with stakeholders for feedback based on initial results from developers
\end{enumerate}

\subsubsection*{2017}

Bluefin session
\begin{enumerate}
\item Review of further results from developers
\item Development of final recommendations to Commission on MP to adopt, together with its associated data inputs.
\end{enumerate}

